Paradigms in Physics 2.0

### 2017-06-29 paper.tex
more edits to vertical figure
everyone's comments accounted for

The resulting curriculum was primarily composed of intensive junior-year \emph{Paradigm}
courses, followed by more conventional senior-year courses.
The \emph{Paradigm} courses meet every day for a total of seven
hours per week for 3 weeks.  These courses incorporate
laboratory experiences and active engagement into the class, and
typically have two problem sets per week.  These courses also had
integrated math content, and were followed by a \emph{Math Methods} course.
The senior-year courses are more traditional 3-credit courses which meet
three hours per week for an entire quarter.

@@

\subsection{Motivation for Paradigms 2.0}
There were a number of factors that motivated us to embark on the
Paradigms~2.0 process.  In the last two decades, we have observed
a number of challenges students face in our major.  While in many cases we
addressed these by changing and reordering the courses, we felt a look
at the entire curriculum was in order.  \davidsays{Triangulation of different sources.  Where did we get info about this issue?}
In addition, our faculty
understanding of the details of the sequence was diminishing by attrition: 
while our younger
faculty are enthusiastic about the Paradigms program, many lack perspective on
where students are in a given course, and what content is essential
for students to grasp prior to a following course.

\paragraph{Transfer students}
Our curriculum is strongly affected by our desire to accommodate
transfer students from community colleges.  This leads us to focus on
a physics major in which all upper-division courses are taken in two
years.  However, our current curriculum is very hard on students in
the fall of their junior year, and especially so for transfer
students.  So we aimed to structure and order the courses such that non-transfer
students could be better prepared for their junior year, while
transfer students could postpone a few courses so that their junior
Fall would be no more difficult than that of any other student.

\paragraph{Coupled curricular changes}
A backlog of curricular changes have accumulated that could not be
separated from an examination of the curriculum as a whole.  Most
notable were discussions of the requirements for computational
physics, the amount of required electronics credits, and the role of
our modern physics course.  These persistent issues were not easily
addressed without addressing the physics major curriculum as a whole,
everything, without progress being made.  
\corinnesays{This paragraph could perhaps be tightened.}

@@

% David R, Ethan, Oksana, Bo, Guenter, Weihong, Matt, Yun-Shik, Heidi,
% Liz, Davide = 11
Only 5 of our 16 current tenure-line faculty were involved in the
orignal Paradigms~1.0 effort.  Some of the newest faculty have only a
superficial understanding of our course structure.  This poses a
challenge when these professors teach courses that are closely
intertwined with one another.

@@

\subsection{5-week \emph{Paradigm} courses}
\davidsays{Advising flexibility, student focus, cohort building---daily,
active engagement.}
To begin with the most distinctive feature of our curriculum, we
wanted to maintain the existing intensive 7-hour-per-week schedule in
the junior year, which we have found effective.  However, we chose to
change from three 3-week courses per quarter to two 5-week courses
per quarter.    This change has several advantages.  \davidsays{Allows opportunity for
formative assessment and feedback for students.  Gives a chance for students to
catch up who fall behind.  Time to adjust to a new professor.  Simplifies teaching
loads} It gives students
a bit more time with each professor prior to their final exam.  This
gives a bit more flexibility, for instance, in case of illness.
Finally, by reducing the number of courses, we allow the curriculum
to be taught with fewer faculty, albeit at an increased work load per professor.

@@

A final change to Electronics is that we now will \emph{require}
\emph{Electronics} during the junior year (specifically as a prerequisite for
\emph{Oscillations and Waves}).  This is made possible by the reduction in
Fall workload for incoming transfer students, who had previously
usually taken \emph{Electronics} as a senior.  This change has enabled us to
articulate distinct and sequenced learning outcomes from these two
courses, particularly in the realm of complex exponentials and Fourier
transforms.  It is also beneficial in providing students with
trouble-shooting skills prior to the in-class labs taught in
\emph{Oscillations and Waves}.

@@

\subsection{Sophomore courses}
We introduced two new sophomore-level courses: \emph{Physics of
Contemporary Challenges} and \emph{Techniques of Theoretical Mechanics}.  Both of
these courses ramp up student mathematical abilities prior to their
junior year.  The \emph{Challenges} course focuses on estimation,
dimensional reasoning, and interpretation of integrals, while
\emph{Theoretical Mechanics} teaches power series approximations and
exposes students to increased levels of mathematical sophistication. \davidsays{ADD CITATIONS SOMEWHERE HERE TO KELBY AND MACKENZIE}

@@

In \emph{The Physics of Contemporary Challenges} we prepare students to
apply physics concepts and physical reasoning skills to sustainable
energy issues, climate change mechanisms, space exploration and
puzzles in fundamental physics. These ``real-world'' topics are chosen
for either the societal need (energy and climate), and/or the human
need to explore (space and fundamental physics). By prioritizing
inclusion of the most engaging challenges, we aim to attract and
retain as many potential physics majors as possible.

@@

The \emph{Theoretical Mechanics} course is aimed at students who are taking
or have completed the last course in the introductory sequence. The
physics content of the course is advanced Newtonian mechanics,
introduction to Lagrangian and Hamiltonian techniques, and special
relativity. These topics at this stage for the students are convenient
for explicit discussion of sense-making strategies in physics because
(a) students are have recently studied problems that serve as limiting
cases for these more complex problems, (b) students at this level are
transitioning from solving problems primarily involving numbers to
problems with only symbolic parameters, and (c) sense-making as strategies 
can be discussed for approaching a problem, evaluating
answers, and refining intuitions about relativistic and other
unfamiliar situations. For example, we teach students how to use
spacetime diagrams to develop a story about a relativistic situation
and how to use hyperbola trigonometry on these diagrams to perform Lorentz
transformations that can be checked against the results of algebraic
Lorentz transformations.  The sense-making goals of the course are on
an equal footing with the physics goals, and sense-making is integrated
explicitly in all aspects of the course: exams, homework and during
in-class activities. \davidsays{CITE HERE?} Our aim is that students will develop
sense-making skills that will improve their learning in advanced
courses and will be valued by their upper-division course instructors,
research advisors, and future employers.

@@

Future work will involve documenting the learning trajectories that we
have developed in our curriculum.  Furthermore, we are engaging in a
project to study our students' development of sense-making skills in
particular through the two new courses that are intended specifically
to ramp up those skills (instincts?).

>>>>

The resulting curriculum was primarily composed of intensive junior-year \emph{Paradigm}
courses, followed by more conventional senior-year courses.
The \emph{Paradigm} courses meet every day for a total of seven
hours per week for 3 weeks.  These courses incorporate
laboratory experiences and active engagement into the class, and
typically have two problem sets per week.  These courses also had
integrated math content, and were followed by a \emph{Math Methods} course.
The senior-year courses are more traditional 3-credit courses which meet
three hours per week for an entire 10-week quarter.

@@

\subsection{Motivation for Paradigms 2.0}
There were a number of factors that motivated us to embark on the
Paradigms~2.0 process.  In the last two decades, we have observed
a number of challenges students face in our major.  While in many cases we
addressed these by changing and reordering the courses, we felt a look
at the entire curriculum was in order.  \davidsays{Triangulation of different sources.  Where did we get info about this issue?}
In addition, our faculty
understanding of the details of the sequence was diminishing by attrition: 
while our younger
faculty are enthusiastic about the Paradigms program, many lack perspective on
where students are in a given course, and what content is essential
for students to grasp for a subsequent course.

\paragraph{Transfer students}
Our curriculum is strongly affected by our desire to accommodate
transfer students from community colleges, who make up about a \davidsays{quarter}
of our Physics majors.  This leads us to focus on
a physics major in which all upper-division courses are taken in two
years.  However, our current curriculum is very hard on students in
the fall of their junior year, and especially so for transfer
students.  So we aimed to structure and order the courses such that non-transfer
students could be better prepared for their junior year, while
transfer students could postpone a few courses so that the
Fall quarter of their junior year would be no more difficult than that of any other student.

\paragraph{Coupled curricular changes}
A backlog of curricular changes had accumulated that could not be
separated from an examination of the curriculum as a whole.  The requirements
for computational physics, the number of required electronics credits, and the
role of our modern physics course could not easily be addressed separately.

@@

% David R, Ethan, Oksana, Bo, Guenter, Weihong, Matt, Yun-Shik, Heidi,
% Liz, Davide = 11
Only 5 of our 16 current tenure-line faculty were involved in the
original Paradigms~1.0 effort.  Some of the newest faculty have only a
superficial understanding of our course structure.  This poses a
challenge when these professors teach courses that are closely
intertwined with one another.

@@

\subsection{5-week \emph{Paradigm} courses}
\davidsays{Advising flexibility, student focus, cohort building---daily,
active engagement.}
To begin with the most distinctive feature of our curriculum, we
wanted to maintain the existing intensive 7-hour-per-week schedule in
the junior year, which we have found effective.  However, we chose to
change from three 3-week courses per quarter to two 5-week courses
per quarter.    This change has several advantages.  \davidsays{Allows opportunity for
formative assessment and feedback for students.  Gives a chance for students to
catch up who fall behind.  Time to adjust to a new professor.  Simplifies teaching
loads} It gives students
a bit more time with each professor prior to their final exam.  This
gives more flexibility, for instance, in case of illness.
Finally, by reducing the number of courses, we allow the curriculum
to be taught with fewer faculty, albeit at an increased work load per professor.

@@

A final change to Electronics is that we now will \emph{require}
\emph{Electronics} during the junior year (specifically as a prerequisite for
\emph{Oscillations and Waves}).  This is made possible by the reduction in
Fall workload for incoming transfer students, who had
usually taken \emph{Electronics} as a senior.  This change has enabled us to
articulate distinct and sequenced learning outcomes from these two
courses, particularly in the realm of complex exponentials and Fourier
transforms.  It is also beneficial in providing students with
trouble-shooting skills prior to the in-class labs taught in
\emph{Oscillations and Waves}.

@@

\subsection{Sophomore courses}
We introduced two new sophomore-level courses: \emph{Physics of
Contemporary Challenges} and \emph{Techniques of Theoretical Mechanics}.  Both of
these courses ramp up student mathematical abilities prior to their
junior year.  The \emph{Challenges} course focuses on estimation,
dimensional reasoning, and interpretation of integrals, while
\emph{Theoretical Mechanics} teaches power series approximations and
exposes students to increased levels of mathematical sophistication and
sense-making strategies. \davidsays{ADD CITATIONS SOMEWHERE HERE TO KELBY AND MACKENZIE}

@@

In \emph{The Physics of Contemporary Challenges} we prepare students to
apply physics concepts and physical reasoning skills to sustainable
energy issues, climate change mechanisms, space exploration and
puzzles in fundamental physics. These ``real-world'' topics are chosen
for either the societal need (energy and climate), and/or the human
need to explore (space and fundamental physics). By prioritizing
inclusion of the most engaging challenges, we aim to attract and
retain as many potential physics majors as possible.
\davidsays{Add in reference and note?}

@@

The \emph{Theoretical Mechanics} course is aimed at students who are taking
or have completed the last course in the introductory sequence. The
physics content of the course is advanced Newtonian mechanics,
introduction to Lagrangian and Hamiltonian techniques, and special
relativity. These topics are convenient
for explicit discussion of sense-making strategies in physics because
(a) these students have recently studied problems that serve as limiting
cases for these more complex problems, (b) students at this level are
transitioning from solving problems primarily involving numbers to
problems with only symbolic parameters, and (c) sense-making as strategies 
can be discussed for approaching a problem, evaluating
answers, and refining intuitions about relativistic and other
unfamiliar situations. For example, we teach students how to use
spacetime diagrams to develop a story about a relativistic situation
and how to use hyperbola trigonometry on these diagrams to perform Lorentz
transformations that can be checked against the results of algebraic
Lorentz transformations.  The sense-making goals of the course are on
an equal footing with the physics goals, and sense-making is integrated
explicitly in all aspects of the course: exams, homework and during
in-class activities. \davidsays{CITE HERE?} Our aim is that students will develop
sense-making skills that will improve their learning in advanced
courses and will be valued by their upper-division course instructors,
research advisors, and future employers.

@@

Future work will involve documenting the learning trajectories that we
have developed in our curriculum.  Furthermore, we are engaging in a
project to study our students' development of sense-making skills in
particular through the two new courses that are intended specifically
to ramp up those skills.
